Rigidity of Expectations: Additional Evidence from Density Forecasts of Professionals and Households

### 2019-07-11 paper/Draft.tex
finished adding additionl texts about weak efficiency tests

	\begin{abstract}
		Surveys on density forecasts of macroeconomic variable that are not availables until recently provide one additional moment restriction, uncertainty, for testing and exploring the implications of various theories on expectation formation. This paper first documents the persistent dispersion in uncertainty about future inflation as well as other stylized facts across economists and households.  Second,  utilizing the panel data of individual uncertainty, I provide additional test results and estimates that are not consistent with the benchmark prediction of full-information rational expectation (FIRE). Finally, using externally identified shocks to inflation, I show the dynamic responses of forecasting moments to shocks are broadly consistent with theories of expectation rigidity and additional evidence for shock-dependent nature of the rigidity.  This sets the stage for estimating shock-specific rigidity using uncertainty as the next step of the analysis.

@@

	Third, \citet{manski2004measuring} and many other papers have advocated long for eliciting probabilistic questions measuring subjective uncertainty in economic surveys. Although the initial suspicion concerning to people's ability in understanding, using and answering probabilistic questions is understandable, \citet{bertrand2001people} and other work have all shown that this is actually less of a concern. \citet{armantier2017overview} have a thorough discussion on designing, experimenting and implementing the consumer expectation surveys to insure the quality of the responses. Broadly speaking, the advocators have argued that going beyond the revealed preference approach, availability to survey data provides economists with direct information on agents' expectations and helps avoids imposing arbitrary assumption.  This insight holds for not only point forecast, but also and even more importantly, for uncertainty.

@@

	Throughout the paper, I use three measures of inflation: headline CPI, core CPI and core PCE . Depending on the specific variable of forecast in the survey series, the realization of the corresponding inflation is used to compute moments such as forecast errors. Specifically, SPF has density forecasts for both core CPI and core PCE \footnote{SPF also  has density forecasts for GDP deflator(for GNP prior to 1992:Q1) going back to 1968. Since the ranges of the values as we as the definition have not been consistent over time. I do not use it in this paper.}. For SCE, as the households responders are asked about the overall inflation, It is the most appropriate to be interpreted as  headline CPI inflation. To simplify the expression, from now on, core CPI and core PCE are simply referred to as CPI and PCE, respectively.

@@

	Another pattern worth discussing in Figure \ref{IQR_Unceratitny} is that there is a notable rise in the dispersion of professional forecasts in the recent 2-3 years, primarily driven by an increase of upper side of the forecast (i.e. 75 percentile forecast increases from 1\% to 2\%). \footnote{This should be interpreted with caution since the disagreements of SPF forecasts shown in bottom row of Figure \ref{UnceratitnyOtherMoments} actually exhibits a gradual decline. } This is consistent with the observation in the top left two graphs of Figure \ref{Unceratitny_Histogram} that the distribution of inflation forecasts in recent years have become flattened.

@@

	\begin{figure}[ht]
		\centering
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Realized Inflation and Uncertainty}
			\label{InfVar}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPICore_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_PCECore_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPIAU_varSCEM.png}
		\end{subfigure}
	    \vspace{1em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Size of Forecast Errors and Uncertainty}
			\label{FEVar}
			\includegraphics[width=0.3\textwidth]{figures/SPFCPI_abFE_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SPFPCE_abFE_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SCE_abFE_varSCEM.png}
		\end{subfigure}
	 \vspace{0.5em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Disagreement and Uncertainty}
			\label{DisgVar}
			\includegraphics[width=0.3\textwidth]{figures/CORECPI_disg_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/COREPCE_disg_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Q9_disg_varSCEM.png}\\
		\end{subfigure}
		\\
		\begin{flushleft}
			{\footnotesize Note:  From left to right: SPF's  forecasts of core CPI and core PCE, and SCE's forecast of headline CPI. From top to the bottom, uncertainty (red dash) versus realized inflation (blue dot) with correlation coefficient of 0.14, 0.12 and -0.25, respectively; uncertainty (red dash) versus absolute value of forecast errors (blue dot) with correlation coefficient of -0.19, -0.18, 0.27, respectively; uncertainty (red dash) versus  disagreements (blue dot) with correlation coefficient of -0.31, -0.37 and 0.43,repectively. Only the Pearson tests of correlation between disagreement and uncertainty are significant for all.}
		\end{flushleft}
		\caption{Uncertainty and Other Moments}
		\label{UnceratitnyOtherMoments}
	\end{figure}

@@

	\begin{figure}[ht]
		\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\caption{Revision in Mean Forecasts}
		\includegraphics[width=7cm]{figures/PRCCPIMean1_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEMean1_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/SCEmean_hist.png} 
		\end{subfigure}
	   \hfill 
	  	\begin{subfigure}[b]{0.5\textwidth}
	  	\caption{Revision in Uncertainty}
		\includegraphics[width=7cm]{figures/PRCCPIVar1_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEVar1_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/SCEvar_hist.png}  
		\end{subfigure}
		\caption{Distribution of Mean Forecast and Uncertainty }
		\label{Unceratitny_Histogram}
	\end{figure}

	\begin{figure}[ht]
		\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=7cm]{figures/PRCCPIMean01_rv_true_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEMean01_rv_true_hist.png} 
		\end{subfigure}
		   \hfill 
		\begin{subfigure}[b]{0.5\textwidth}
		\includegraphics[width=7cm]{figures/PRCCPIVar01_rv_true_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEVar01_rv_true_hist.png} 
		\end{subfigure}
		\caption{Distribution of Revision in Forecasts and Uncertainty}
		\label{RevisionHist}
	\end{figure}

@@

	This section first reproduces a number of statistical tests of FIRE used seen existing literature in Table \ref{NullTestTable}, primarily following \citet{mankiw2003disagreement},  and then extends the tests relying on uncertainty in Table \ref{RevEfficiency} and \ref{WeakRevEfficiency}, in the spirit of forecasting efficiency by \citet{nordhaus1987forecasting}. It is an extension of revision tests on mean forecasts by \citet{fuhrer2018intrinsic} to uncertainty.

@@

	Second, the point forecast one year ago predicts the forecast errors for professionals in the significance level of 1\%. For headline CPI inflation, for instance, one percentage point inflation forecast corresponds to 0.3 percentage points of the forecast errors one year later. Thus test 2 in Table \ref{NullTestTable} easily rejects the second hypothesis test of FIRE that past information does not predict future forecast errors.

@@

	Table \ref{WeakRevEfficiency} presents the results with the revision replaced with change in mean forecasts and uncertainty, i.e. $y_{t|t-1}$ to $y_{t+1|t}$. The auto-correlation of change in mean forecast and uncertainty does not embed tests of null of FIRE (Equation \ref{VarREPop}). But if the individual forecasts and their uncertainty are persistent over different horizons, this implies that the agent does not react to the news and newly realized shocks enough.

@@

@@

	I use three types of shocks: technology(productivity) shock from \citet{gali1999technology}, oil price shock from \citet{hamilton1996happened} and monetary policy shocks from  \citet{laseen2011anticipated}. The first two are the same in \citet{coibion2012can}. \footnote{ Another shock in \citet{coibon2012can} is the news shock by \citet{barsky2011news}. This will be included in later version of the paper.}

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[width=0.7\textwidth]{figures/inf_shocksQ.png}
			\begin{flushleft}
			{\footnotesize Note: all the shocks are normalized by their sample standard deviation. Technology shock is idetified by long-run restriction of structural VAR with labor productivity, hours and inflation. Oil shock is defined as the max \{0, the largest increase of oil price over the past 12 months.\} }
		\end{flushleft}
		\caption{ Shocks to Inflation}
		\label{Inflationshocks}
	\end{figure}

@@

	As to the households forecast, the starkly wider confidence band across  shocks and moments seen in Figure \ref{SCE_IR} are not that surprising given the wide dispersion of the forecasts moments of a dramatically larger magnitude seen in Figure \ref{UnceratitnyOtherMoments}.

>>>>

	\begin{abstract}
		Surveys on density forecasts of macroeconomic variable that are not available until recently provide one additional moment restriction, uncertainty, for testing and exploring the implications of various theories on expectation formation. This paper first documents the persistent dispersion in uncertainty about future inflation as well as other stylized facts across economists and households.  Second,  utilizing the panel data of individual uncertainty, I provide additional test results and estimates that are not consistent with the benchmark prediction of full-information rational expectation (FIRE). Finally, using externally identified shocks to inflation, I show the dynamic responses of forecasting moments to shocks are broadly consistent with theories of expectation rigidity and additional evidence for shock-dependent nature of the rigidity.  This sets the stage for estimating shock-specific rigidity using uncertainty as the next step of the analysis.

@@

	Third, \citet{manski2004measuring},\citet{delavande2011measuring} and many other papers have advocated long for eliciting probabilistic questions measuring subjective uncertainty in economic surveys. Although the initial suspicion concerning to people's ability in understanding, using and answering probabilistic questions is understandable, \citet{bertrand2001people} and other work have shown respondents’ consistent ability and willingness to assign a probability (or “percent chance”) to future events. \citet{armantier2017overview} have a thorough discussion on designing, experimenting and implementing the consumer expectation surveys to insure the quality of the responses\footnote{Other literature includes \citet{van2008rethinking},\citet{delavande2014probabilistic}. etc.}. Broadly speaking, the advocators have argued that going beyond the revealed preference approach, availability to survey data provides economists with direct information on agents' expectations and helps avoids imposing arbitrary assumption.  This insight holds for not only point forecast, but also and even more importantly, for uncertainty.

@@

	Throughout the paper, I use three measures of inflation: headline CPI, core CPI and core PCE . Depending on the specific variable of forecast in the survey series, the realization of the corresponding inflation is used to compute moments such as forecast errors. Specifically, SPF has density forecasts for both core CPI and core PCE \footnote{SPF also  has density forecasts for GDP deflator(for GNP prior to 1992:Q1) going back to 1968. Since the ranges of the values and the definition are not consistent over time, I do not use it in this paper.}. For SCE, as the households responders are asked about the overall inflation, It is the most appropriate to be interpreted as  headline CPI inflation. To simplify the expression, from now on, core CPI and core PCE are simply referred to as CPI and PCE, respectively.

@@

	Another pattern worth discussing in Figure \ref{IQR_Unceratitny} is that there is a notable rise in the dispersion of professional forecasts in the recent 2-3 years, primarily driven by an increase of upper side of the forecast (i.e. 75 percentile forecast increases from 1\% to 2\%). \footnote{This should be interpreted with caution since the disagreements of SPF forecasts shown in bottom row of Figure \ref{UncertaintyOtherMoments} actually exhibits a gradual decline. } This is consistent with the observation in the top left two graphs of Figure \ref{Unceratitny_Histogram} that the distribution of inflation forecasts in recent years have become flattened.

@@

	\begin{figure}[ht]
		\centering
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Realized Inflation and Uncertainty}
			\label{InfVar}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPICore_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_PCECore_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Inf1yf_CPIAU_varSCEM.png}
		\end{subfigure}
	    \vspace{1em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Size of Forecast Errors and Uncertainty}
			\label{FEVar}
			\includegraphics[width=0.3\textwidth]{figures/SPFCPI_abFE_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SPFPCE_abFE_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/SCE_abFE_varSCEM.png}
		\end{subfigure}
	 \vspace{0.5em}
		\vfill
		\begin{subfigure}[b]{\textwidth}
			\centering
			\caption{Disagreement and Uncertainty}
			\label{DisgVar}
			\includegraphics[width=0.3\textwidth]{figures/CORECPI_disg_varSPFCPIQ.png}
			\includegraphics[width=0.3\textwidth]{figures/COREPCE_disg_varSPFPCEQ.png}
			\includegraphics[width=0.3\textwidth]{figures/Q9_disg_varSCEM.png}\\
		\end{subfigure}
		\\
		\begin{flushleft}
			{\footnotesize Note:  From left to right: SPF's  forecasts of core CPI and core PCE, and SCE's forecast of headline CPI. From top to the bottom, uncertainty (red dash) versus realized inflation (blue dot) with correlation coefficient of 0.14, 0.12 and -0.25, respectively; uncertainty (red dash) versus absolute value of forecast errors (blue dot) with correlation coefficient of -0.19, -0.18, 0.27, respectively; uncertainty (red dash) versus  disagreements (blue dot) with correlation coefficient of -0.31, -0.37 and 0.43,repectively. Only the Pearson tests of correlation between disagreement and uncertainty are significant for all.}
		\end{flushleft}
		\caption{Uncertainty and Other Moments}
		\label{UncertaintyOtherMoments}
	\end{figure}

@@

	\begin{figure}[ht]
		\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\caption{Mean Forecasts}
		\includegraphics[width=7cm]{figures/PRCCPIMean1_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEMean1_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/SCEmean_hist.png} 
		\end{subfigure}
	   \hfill 
	  	\begin{subfigure}[b]{0.5\textwidth}
	  	\caption{Uncertainty}
		\includegraphics[width=7cm]{figures/PRCCPIVar1_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEVar1_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/SCEvar_hist.png}  
		\end{subfigure}
		\caption{Distribution of Mean Forecast and Uncertainty }
		\label{Unceratitny_Histogram}
	\end{figure}

	\begin{figure}[ht]
		\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\caption{Revision in Mean Forecasts}
		\includegraphics[width=7cm]{figures/PRCCPIMean01_rv_true_hist.png} 
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEMean01_rv_true_hist.png} 
		\end{subfigure}
		   \hfill 
		\begin{subfigure}[b]{0.5\textwidth}
					\caption{Revision in Uncertainty}
		\includegraphics[width=7cm]{figures/PRCCPIVar01_rv_true_hist.png}  
		\smallskip
		\includegraphics[width=7cm]{figures/PRCPCEVar01_rv_true_hist.png} 
		\end{subfigure}
		\caption{Distribution of Revision in Forecasts and Uncertainty}
		\label{RevisionHist}
	\end{figure}

@@

	This section first reproduces a number of statistical tests of FIRE seen in existing literature in Table \ref{NullTestTable}, primarily following \citet{mankiw2003disagreement},  and then extends the tests relying on uncertainty in Table \ref{RevEfficiency} and \ref{WeakRevEfficiency}, in the spirit of forecasting efficiency by \citet{nordhaus1987forecasting}. It is an extension of revision tests on mean forecasts by \citet{fuhrer2018intrinsic} to uncertainty.

@@

	Second, the point forecast one year ago predicts the forecast errors for professionals in the significance level of $1\%$. For headline CPI inflation, for instance, one percentage point inflation forecast corresponds to 0.3 percentage points of the forecast errors one year later. Thus test 2 in Table \ref{NullTestTable} easily rejects the second hypothesis test of FIRE that past information does not predict future forecast errors.

@@

	Table \ref{WeakRevEfficiency} presents the results with the revision replaced with change in mean forecasts and uncertainty, i.e. from $y_{t|t-1}$ to $y_{t+1|t}$. As we have discussed in Section \ref{theory}, the auto-correlation of change in mean forecast and uncertainty do not bear testable predictions from FIRE (Equation \ref{VarREPop}). But if the  forecasts and uncertainty are persistent in its first difference, it may imply that the agent does not react to the news and newly realized shocks sufficiently enough. In addition, the auto-correlation regressions of this kind is a useful characterization of the time series dynamics of forecasts. With the variable being the first difference, the panel structure of SCE and SPF allows for calculating changes in individual levels for a greater sample size, especially for households. Besides autoregression, I also report the constant estimate of the changes in the first column of each sub panel.

    The most noticeable pattern for both professionals and households, and for both mean forecast and uncertainty is that past change predict future changes with universally negative coefficients across different horizons. Most of the negative coefficients are statistically significant in the 1\% level. For instance, one percentage point of increase in SPF's CPI forecast in the previous quarter predicts around 0.29 percentage points decreases in next quarter. This negative correlation are smaller in size but remain significant into further past, i.e. -0.24 for two quarters lag and -0.1 for three quarters lag. Data on SCE is monthly, so lags are included up to six months. The negative correlation between past and current changes are all negative and significant for households. The sizes of the correlation coefficients are comparable with professional forecasts when the monthly coefficients are converted to their quarterly average, i.e. -0.3 to -0.4.

    Such an auto-correlation of change in mean forecast is very much reflected in the same regression for uncertainty, as reported in the bottom panel of Table \ref{WeakRevEfficiency}.  For SPF of CPI and PCE, respectively, one unit increase in uncertainty about 1-year-ahead inflation in previous quarter predicts around a 0.39 and 0.44 unit of the drop in next quarter. The effect holds up to two quarters for professionals and 5 months for households.

    These evidences suggest that both mean forecast and uncertainty of individuals are mean-reverting. An essentially equivalent explanation is that both series are realizations of noisy signals around their respective long-run mean. This will lead to exactly negative correlation of first differences as we have seen. The mean-reverting patterns might also explain what is observed from Figure \ref{Unceratitny_Histogram}, according to which, there is no significant changes in the distribution across different years.

    The second noticeable result lies in the constant regressions reported in the first column of each sub panel in Table \ref{WeakRevEfficiency}. It implies that households constantly lower their mean forecasts as well as uncertainty from month to month, while professional forecasts do not behave in such a way. In particular, the constant regression of change in mean forecast for SCE gives an estimated coefficient of -0.05 which is significant in the $5\%$ level. Individual households' 1-year-ahead inflation expectations keep being downward adjusted each month compared to their previous answer. What is more interesting is that their uncertainty about 1-year-ahead inflation also decrease each month.  The size of downward adjustment is  $-1.39$ unit and statistically significant in the level of $1\%.$. This negative significant constant coefficient remains throughout all auto-regressions, implying it is not driven by time-varying changes.

   The most natural explanation for this that repeatedly surveyed households have become more informative about inflation over time. Given the unconditional forecast errors of inflation by households are positive, a downward adjustment of inflation stands for a less-biased forecast.   \footnote{The possibility that the surveys' information set is influenced by the survey itself is a double-edged sword. One one hand, this poses a methodological challenge to the survey designers of expectations as to if surveys can objectively elicit the ``true’’ expectations held by the respondents. On the other hand, researchers can use survey as a meaningful intervention tool to identify the effect of factors such as information provision and attention. Recent examples of this type of research includes: \citet{coibion2018firms} for firms and \citet{coibion2019monetary} for households.}

@@

  In summary, the major additional insights that arise from the empirical tests of this section is that rigidity of incorporating new information in forming expectations imply noticeable inefficiency of revisions in forecasts and drop in uncertainty.

@@

	I use three types of shocks: technology(productivity) shock from \citet{gali1999technology}, oil price shock from \citet{hamilton1996happened} and monetary policy shocks from  \citet{laseen2011anticipated}. The first two are the same in \citet{coibion2012can}. \footnote{ \citet{coibion2012can} also uses the news shock by \citet{barsky2011news}. This will be included in later version of the paper.}

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[width=0.8\textwidth,totalheight=0.6\textwidth]{figures/inf_shocksQ.png}
			\begin{flushleft}
			{\footnotesize Note: all the shocks are normalized by their sample standard deviation. Technology shock is idetified by long-run restriction of structural VAR with labor productivity, hours and inflation. Oil shock is defined as the max \{0, the largest increase of oil price over the past 12 months.\} }
		\end{flushleft}
		\caption{ Shocks to Inflation}
		\label{Inflationshocks}
	\end{figure}

@@

	As to the households forecast, the starkly wider confidence band across  shocks and moments seen in Figure \ref{SCE_IR} are not that surprising given the wide dispersion of the forecasts moments of a dramatically larger magnitude seen in Figure \ref{UncertaintyOtherMoments}.
